Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2015-08-28 doc/GMP2015/revision_3/s01_introduction.tex
Made a talk, consolidated doc/img, began new revision, updated doc todo and .gitignore
Updated paper to say algo is O(V^4). However, I also provide a sketch of what I believe to be a working O(V^3). Updated todo's.
Lots of work on Revision 3, getting close.
Revision 3, close to finished.

Na{\"i}vely, the optimal DR-plan is used as follows. Each decomposed subsystem---a node of the DR-plan---is treated and solved as a polynomial system of constraints between its child subsystems. However, even in an optimal DR-plan, there can be arbitrarily many children at a node. In other words, even in the recursive decomposition given by an optimal DR-plan, the  size of the maximal indecomposable subsystem could be arbitrarily large.  It represents a bottleneck that dictates the complexity of solving or realizing the constraint system \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}.  We address this problem using the recently developed concept of \dfn{convex Cayley configuration spaces} \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII,sitharam2014beast,sitharam2013caymos,wang2014cayley}. This allows for even greater reduction of the complexity by realizing large, indecomposable systems in a manner that avoids working with large systems of equations.
Specifically, we give an efficient technique for \dfn{optimally modifying} large indecomposable subsystems in a manner that reduces their complexity while preserving desired solutions; the modification ensures a convex Cayley configuration space, and the space can be efficiently searched to find a realization that satisfies the additional constraints of the original system.
This optimal modification  problem is a generalization of the previously studied problem of optimal completion of underconstrained systems \cite{sitharam2005combinatorial,joan-arinyo2003transforming}.

DR-plans are especially useful for constraint systems that exhibit some level of \dfn{self-similarity} and \dfn{quasi-uniformity}, in addition to isostaticity.  These properties can be leveraged to further reduce the complexity of both optimal DR-plan construction and recombination.
We consider 3 different types of constraint systems---which we collectively call \dfn{qusecs}---that are used to model, design, and analyze quasi-uniform or self-similar materials.
In the remainder of this section, we motivate the materials application and give the contributions and organization of the paper.

@@

% \FigInit{}{fig:material_examples}
% \FigThreeSubfig%
%   {img/Chlamydomonas_TEM_17}
%   {Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}
%   {fig:material_examples:microtubule}%
%   %
%   {img/ligten2}
%   {Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}
%   {fig:material_examples:tendon}%
%   %
%   {img/Rothemund-DNA-SierpinskiGasket}
%   {A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}
%   {fig:material_examples:sierpinski}

\ClearMyMinHeight
\SetMyMinHeight{.32}{img/Chlamydomonas_TEM_17}
\SetMyMinHeight{.32}{img/ligten2}
\SetMyMinHeight{.32}{img/Rothemund-DNA-SierpinskiGasket}

\begin{figure*}\centering%
  %
  \begin{subfigure}{0.32\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/Chlamydomonas_TEM_17}
    \caption{}\label{fig:material_examples:microtubule}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.32\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/ligten2}
    \caption{}\label{fig:material_examples:tendon}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.32\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/Rothemund-DNA-SierpinskiGasket}
    \caption{}\label{fig:material_examples:sierpinski}
  \end{subfigure}%
  %
  \caption{(\ref{fig:material_examples:microtubule}) Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}. (\ref{fig:material_examples:tendon}) Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}. (\ref{fig:material_examples:sierpinski}) A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}\label{fig:material_examples}
\end{figure*}%

% \FigAddSubfig%
%   {0.3}
%   {img/Chlamydomonas_TEM_17}
%   {Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}
%   {fig:material_examples:microtubule}%
% \FigAddSubfig%
%   {0.3}
%   {img/ligten2}
%   {Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}
%   {fig:material_examples:tendon}%
% \FigAddSubfig%
%   {0.3}
%   {img/Rothemund-DNA-SierpinskiGasket}
%   {A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}
%   {fig:material_examples:sierpinski}
% \FigDisplay{}{fig:material_examples}

@@

% Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}).
% The canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems.
% Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied Cluster Minimality \cite{hoffman2001decompositionI} (see Figure~\ref{fig:demo_graph:clustmindrp}).

In Section \ref{sec:DRP}, we define a so-called \dfn{canonical DR-plan} and prove a strong Church-Rosser property: all canonical DR-plans for isostatic or underconstrained qusecs are optimal. In so doing, we navigate the NP-hardness barrier present in the general form of the DR-planning problem; the canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans when a system is over-constrained. Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied \dfn{cluster minimality}~\cite{hoffman2001decompositionI}. Also in this section, a polynomial time (\candrpcomplexity) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.  While this and the next section focus on bar-joint graphs, the theory is easily extended to other qusecs used to model the abovementioned types of materials, as shown in subsequent sections.
% that such a plan will be optimal, in the sense that it minimizes the fan-in over all nodes, for isostatic bar-joint graphs.
% Additionally, a polynomial time (\candrpcomplexity) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.

>>>>

Na{\"i}vely, the optimal DR-plan is used as follows. Each decomposed subsystem -- a node of the DR-plan -- is treated and solved as a polynomial system of constraints between its child subsystems. However, even in an optimal DR-plan, there can be arbitrarily many children at a node. In other words, even in the recursive decomposition given by an optimal DR-plan, the  size of the maximal indecomposable subsystem could be arbitrarily large.  It represents a bottleneck that dictates the complexity of solving or realizing the constraint system \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}.  We address this problem using the recently developed concept of \dfn{convex Cayley configuration spaces} \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII,sitharam2014beast,sitharam2013caymos,wang2014cayley}. This allows for even greater reduction of the complexity by realizing large, indecomposable systems in a manner that avoids working with large systems of equations.
Specifically, we give an efficient technique for \dfn{optimally modifying} large indecomposable subsystems in a manner that reduces their complexity while preserving desired solutions; the modification ensures a convex Cayley configuration space, and the space can be efficiently searched to find a realization that satisfies the additional constraints of the original system.
This optimal modification  problem is a generalization of the previously studied problem of optimal completion of underconstrained systems \cite{sitharam2005combinatorial,joan-arinyo2003transforming}.

DR-plans are especially useful for constraint systems that exhibit some level of \dfn{self-similarity} and \dfn{quasi-uniformity}, in addition to isostaticity.  These properties can be leveraged to further reduce the complexity of both optimal DR-plan construction and recombination.
We consider 3 different types of constraint systems -- which we collectively call \dfn{qusecs} -- that are used to model, design, and analyze quasi-uniform or self-similar materials.
In the remainder of this section, we motivate the materials application and give the contributions and organization of the paper.

@@

% \FigInit{}{fig:material_examples}
% \FigThreeSubfig%
%   {../../img/Chlamydomonas_TEM_17}
%   {Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}
%   {fig:material_examples:microtubule}%
%   %
%   {../../img/ligten2}
%   {Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}
%   {fig:material_examples:tendon}%
%   %
%   {../../img/Rothemund-DNA-SierpinskiGasket}
%   {A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}
%   {fig:material_examples:sierpinski}

\ClearMyMinHeight
\SetMyMinHeight{.32}{../../img/Chlamydomonas_TEM_17}
\SetMyMinHeight{.32}{../../img/ligten2}
\SetMyMinHeight{.32}{../../img/Rothemund-DNA-SierpinskiGasket}

\begin{figure*}\centering%
  %
  \begin{subfigure}{0.32\linewidth}\centering
    \includegraphics[height=\myMinHeight]{../../img/Chlamydomonas_TEM_17}
    \caption{}\label{fig:material_examples:microtubule}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.32\linewidth}\centering
    \includegraphics[height=\myMinHeight]{../../img/ligten2}
    \caption{}\label{fig:material_examples:tendon}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.32\linewidth}\centering
    \includegraphics[height=\myMinHeight]{../../img/Rothemund-DNA-SierpinskiGasket}
    \caption{}\label{fig:material_examples:sierpinski}
  \end{subfigure}%
  %
  \caption{(\ref{fig:material_examples:microtubule}) Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}. (\ref{fig:material_examples:tendon}) Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}. (\ref{fig:material_examples:sierpinski}) A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}\label{fig:material_examples}
\end{figure*}%

% \FigAddSubfig%
%   {0.3}
%   {../../img/Chlamydomonas_TEM_17}
%   {Cross section of the Chlamydomonas algae axoneme, a cilia composed of microtubules~\cite{wikimediacommons2007cilia}.}
%   {fig:material_examples:microtubule}%
% \FigAddSubfig%
%   {0.3}
%   {../../img/ligten2}
%   {Cross section of a tendon displaying the hierarchical structure~\cite{lecture_biosolid_mechanics}.}
%   {fig:material_examples:tendon}%
% \FigAddSubfig%
%   {0.3}
%   {../../img/Rothemund-DNA-SierpinskiGasket}
%   {A DNA array exhibiting the Sierpinski triangle~\cite{wikimediacommons2007dna}.}
%   {fig:material_examples:sierpinski}
% \FigDisplay{}{fig:material_examples}

@@

% Also in Section \ref{sec:DRP}, we give an efficient (\ComplexityCanDRP) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}).
% The canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems.
% Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied Cluster Minimality \cite{hoffman2001decompositionI} (see Figure~\ref{fig:demo_graph:clustmindrp}).

In Section \ref{sec:DRP}, we define a so-called \dfn{canonical DR-plan} and prove a strong Church-Rosser property: all canonical DR-plans for isostatic or underconstrained qusecs are optimal. In so doing, we navigate the NP-hardness barrier present in the general form of the DR-planning problem; the canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans when a system is over-constrained. Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied \dfn{cluster minimality}~\cite{hoffman2001decompositionI}. Also in this section, an \ComplexityCanDRP\ time algorithm is provided to find an optimal DR-plan for independent bar-joint graphs. While this and the next section focus on bar-joint graphs, the theory is easily extended to other qusecs used to model the abovementioned types of materials, as shown in subsequent sections.
% that such a plan will be optimal, in the sense that it minimizes the fan-in over all nodes, for isostatic bar-joint graphs.
% Additionally, a polynomial time (\ComplexityCanDRP) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.

### 2015-08-28 doc/GMP2015/revision_3/s02_preliminaries.tex
Revision 3, close to finished.

%
% \todo{This was labeled as both begin(remark) and end(observation), what was your intent?}
\begin{remark}
  More than one node (leaf) in a DR-plan forest may represent the same subgraph (vertex) of $G$.
  For a given graph, there could be exponentially many DR-plans---and even optimal DR-plans---in the size of the graph. A complete DR-plan is unique but may not be (and is usually not) optimal. DR-plans of self-similar graphs are self-similar.
\end{remark}

See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \todo{we lost this figure, replaced with \ref{fig:seqdrp_nonuniqueness}, should it be cited here now or should the other be added back in somewhere?} \ref{fig:overconstrained}, and \ref{fig:bodypindrp} for examples of DR-plans and how their properties relate to each other.

>>>>

%
% \todo{This was labeled as both begin(remark) and end(observation), what was your intent?}
\begin{remark}
  More than one node (leaf) in a DR-plan forest may represent the same subgraph (vertex) of $G$.
  For a given graph, there could be exponentially many DR-plans -- and even optimal DR-plans -- in the size of the graph. A complete DR-plan is unique but may not be (and is usually not) optimal. DR-plans of self-similar graphs are self-similar.
\end{remark}

See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:overconstrained}, and \ref{fig:bodypindrp} for examples of DR-plans and how their properties relate to each other.
